\Title

### 2021-08-26 latex/paper/main.tex
Full edit of the proof. Needs one or two more passes, and we need to remove over one column of material.

Here is fundamental idea behind this proof:
for each of \(\mathcal{O}_1\) or \(\mathcal{O}_2(\rho)\),
we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
coordinates adapted to level sets of \(E(q_u,p_u)\);
that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
point on that level set.
If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
that \(r\) increases ``on average", then \(E\) must also increase ``on average",
and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
must be gaining energy on the domain of the diffeomorphism.

The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
under which energy will increase on a domain \(D\); later in this section, we
will assign \(D = \mathcal{O}_1\) or \(D = \mathcal{O}_2(\rho)\), find an
appropriate transformation \(T\), and apply this lemma to prove the claims in
Theorem \ref{thm:acrobot-energy-stabilization}.

@@

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    \eqref{eqn:acrobot-constrained-dynamics}.
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
         \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
    \(\dot{\theta} = f_\theta(\theta,r,I)\).
    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
    \((\partial g/\partial r)(\theta,r,0) \equiv 0\) and
    \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
    is \(2\pi\)-periodic in \(\theta\).
    Assume also that there exists \(\epsilon > 0\) such that 
    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
%    \[
%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
%        ,
%    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(\dot{\theta} > 0\) when \(I = 0\) and the constrained dynamics 
    in \((r,\theta)\) coordinates are continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, [-I_1,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
                \hat{r}} = 
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    with state \(\hat{r}\).
    Here \(\theta\) represents time,
    prime denotes differentiation by \(\theta\), and
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Note that \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
    \(g(\theta,\hat{r},0) \equiv 0\).
    The function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}\label{eqn:r1-hat-prime}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        b(\theta,r_0)
        %\frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).

    Since \((\partial g/\partial r)(\theta,r_0,0) \equiv 0\), we get
    \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\).
    The solution of \eqref{eqn:r1-hat-prime} can therefore be obtained by
    quadrature, giving
    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
%    \[
%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
%        .
%    \]
%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Since \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align}\label{eqn:poincare-map}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
    Here, \(r_n\) corresponds to the distance along the line segment
    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
    in \((q_u,p_u)\) coordinates must have departed \(D\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
    Then for each \(I \in ]0,I^\star]\), the value 
    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
    \(P(r_0) \geq r_0 + \gamma\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    and all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).
    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).

    Changing the sign of \(I\) changes the direction of all the
    above inequalites, so that \(r_{n} \leq r_0 - n\gamma\). 
    In this case all orbits of \eqref{eqn:r-discrete} flow below
    \(\underline{r}\), i.e., the constrained dynamics lose energy on \(D\).
\end{proof}

\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
To prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}, we adapt
\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
pseudo-polar coordinates (adapted to level sets of \(E\)) on
the punctured open set \(\mathcal{O}_1 \backslash \{(0,0)\} \subset \SxR\).
The resulting coordinate transformation is
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
One can verify that the inverse to this transformation is
\begin{align*}
    q_u &= rc_\theta
    , \\
    p_u &= -\sign{s_\theta}
    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
    ,
\end{align*}
where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
We now verify these pseudo-polar coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.

When \(I = 0\) we have \(\dot{r} = 0\) and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    \dot{\theta} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}\dot{\theta} 
    = \lim\limits_{\theta \to \pi} \dot{\theta}
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
one can verify that \(\dot{\theta} > 0\) when \(I = 0\).
Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\) as required.

Next we compute
\(b(\theta,r_0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |s_\theta| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
and \(2\pi\)-periodic in \(\theta\).
This means that there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
if and only if there exists \(\delta > 0\) such that 
\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\).
Since \(a(\theta,r_0)\) is adimensional, we can numerically compute \(Q(r_0)\)
once for all acrobots.
The plot of \(Q(r_0)\) when \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
Figure \ref{fig:acrobot-Q}.
We see that it is strictly positive and monotonically increasing, with an
asymptote at \(r_0 = \pi\).
Simulations with smaller \(r_0\) result in an infinite integral error due to a
division by zero, yet we believe the \(Q(r_0)\) is in fact positive for all 
\(r_0 \in \, ]0,\pi[\).
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{Qro.png}
    \caption{The plot of \(Q(r_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

Unfortunately there is no \(\delta > 0\) such that \(Q(r_0) \geq \delta\)
everywhere on \(]0,\pi[\), though such a \(\delta\) does exist 
on \(]\omega,\pi[\) for each fixed \(\omega > 0\).
This means all conditions of Lemma \ref{lemma:energy-regulation-D}
are satisfied on the annulus
\[
    D_\omega = \mathcal{O}_1 \backslash 
    \{(q_u,p_u) \in \SxR \mid E(q_u,p_u) \leq E(\omega,0)\}
    .
\]
Therefore, for each \(\omega \in \, ]0,\pi[\) there exists some 
\(I_1(\omega) > 0\) such that the acrobot gains/loses energy everywhere on
\(D_\omega\).
To prove energy gain on all of \(\mathcal{O}_1\), we will show the origin is a
repeller so that all orbits near the origin flow away from the origin into
\(D_\omega\) (for small enough \(\omega\)) and continue gaining energy
thereafter.

Recall the discrete-time pseudo-radius dynamics \eqref{eqn:r-discrete} from the
proof of Lemma \ref{lemma:energy-regulation-D}. 
To prove the origin is a repeller, we need simply
prove the linearization of \eqref{eqn:r-discrete} is greater than \(1\) at 
\(r_0 = 0\). 
This linearization yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
%\begin{equation*}
%    \lim \limits_{r_0 \to 0^+}
%    \pdiff{a}{r}(\theta,r_0) 
%    = -\frac{\sqrt{2}}{2} \left(11\sin(\theta)^2 - 6\right)
%    ,
%\end{equation*}
%which means
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\).
The origin is therefore a repeller, which means there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
Hence, all orbits in \(\mathcal{O}_1\) near the
origin will gain energy up to (and including) \(E(\mu,0)\).

Since there exists \(I_1(\mu)\) so that the acrobot gains energy on \(D_\mu\),
taking \(I^\star = \min\{I_0,I_1(\mu)\}\) means the acrobot will gain energy on
\(\mathcal{O}_1\) for each \(I \in \, ]0,I^\star]\).
If \(I < 0\), then \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium.
Since the acrobot now loses energy on \(D_\mu\), it also loses energy 
on \(\mathcal{O}_1\) for \(I \in [-I^\star,0[\).

We have proved that any acrobot constrained by 
our VNHC \eqref{eqn:acrobot-constraint} gains/loses energy on \(\mathcal{O}_1\).
We must now confirm that the acrobot will begin rotating, 
i.e., that for \(I > 0\) almost every orbit initialized in \(\mathcal{O}_1\)
will reach the line \(|q_u| = \pi\).
Since the definition of energy gain only guarantees that orbits initialized in
\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
time, we must first prove that almost every orbit in \(\mathcal{O}_1\) will
escape \(\mathcal{O}_1\) through its boundary, the level set \(E_\pi\) with
energy \(E(\pi,0)\).

>>>>

Here is fundamental idea behind this proof:
for each of \(\mathcal{O}_1\) and \(\mathcal{O}_2(\rho)\),
we can find a diffeomorphism \(T(q_u,p_u) = (r,\theta)\) into pseudo-polar
coordinates adapted to level sets of \(E(q_u,p_u)\);
that is, when \(I = 0\) in \eqref{eqn:acrobot-constraint}, the pseudo-radius
\(r\) chooses a level set of \(E\) and the pseudo-angle \(\theta\) chooses a
point on that level set.
If the constrained dynamics in \((r,\theta)\) coordinates flow in such a way
that \(r\) increases ``on average", then \(E\) must also increase ``on average",
and the original constrained dynamics \eqref{eqn:acrobot-constrained-dynamics} 
must be gaining energy on the domain of the diffeomorphism.

The following lemma provides a set of conditions on the \((r,\theta)\) dynamics
under which energy will increase on a domain \(D\); later in this section, we
will assign \(D \subset \mathcal{O}_1\) or \(D \subset \mathcal{O}_2(\rho)\),
find the diffeomorphism \(T\), and apply this lemma to prove the claims in
Theorem \ref{thm:acrobot-energy-stabilization}.

@@

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    \eqref{eqn:acrobot-constrained-dynamics}.
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
         \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
    \(\dot{\theta} = f_\theta(\theta,r,I)\).
    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
    Assume also that \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
    is \(2\pi\)-periodic in \(\theta\),
    and that there exists \(\epsilon > 0\) such that 
    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
%    \[
%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
%        ,
%    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(f_\theta(\theta,r,0) > 0\) and \(f_\theta\)
    is continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, [-I_1,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
                \hat{r}} = 
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    with state \(\hat{r}\), where now
    \(\theta\) represents time,
    prime denotes differentiation by \(\theta\), and
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
    \(g(\theta,\hat{r},0) \equiv 0\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}\label{eqn:r1-hat-prime}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        b(\theta,r_0)
        %\frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).

    Since \((\partial g/\partial \hat{r})(\theta,r_0,0) \equiv 0\),
    the solution of \eqref{eqn:r1-hat-prime} can be obtained by
    quadrature, giving
    \(\hat{r}_1(\theta,r_0) = \int_0^\theta b(\sigma,r_0)d\sigma\).
%    \[
%        \hat{r}_1(\theta,r_0) = \int\limits_0^\theta b(\sigma,r_0)d\sigma
%        .
%    \]
%    with \(b(\theta,r_0) := (\partial g/\partial I)(\theta,r_0,0)\). 
    Recall that \(b(\theta,r_0)\) is \(2\pi\)-periodic in
    \(\theta\), so there is a well-defined Poincar\'{e} map describing
    how \(\hat{r}\) changes each time \(\theta\) increases by \(2\pi\).
    This map is 
    \begin{align}\label{eqn:poincare-map}
        P : \, ]\underline{r},\overline{r}[ &\to \mathbb{R}_{\geq 0} \\
        r_0 &\mapsto \hat{r}(2\pi,r_0,I)
        ,
    \end{align}
    which expands to \(P(r_0) = r_0 + IS(r_0) + R(2\pi,r_0,I)\).
    In other words, we can understand the evolution of \(\hat{r}\) by studying
    the dynamics of the discrete-time system
    \begin{equation}\label{eqn:r-discrete}
        r_{n+1} = P(r_n) =  r_n + IS(r_n) + R(2\pi,r_n,I)
        ,
    \end{equation}
    with initial condition \(r_0\).
%    Here, \(r_n\) corresponds to the distance along the line segment
%    \(L := ]\underline{r},\overline{r}[ \, \times \{0\}\)
%    when an orbit of \((r,\theta)\) hits \(L\) for the the \(n^\text{th}\) time,
%    assuming the acrobot was initialized at \((r,\theta) = (r_0,0)\).

    If at any point in discrete time \(r_n\) exits \(]\underline{r},\overline{r}[\), the orbit 
    in \((q_u,p_u)\) coordinates must have departed \(D\).
    Hence, proving the constrained dynamics gain energy on \(D\) is
    equivalent to finding \(N \in \mathbb{Z}_{> 0}\) such that 
    \(r_{n} \geq \overline{r}\) for all \(n \geq N\).
    Using a perturbation analysis (see Khalil 
    \cite[Theorem 10.1]{khalil_nonlinear}),
    we know there exists some \(I_2 \in \, ]0,I_1]\) and \(\kappa > 0\)
    such that \(\left|R(\theta,r,I)\right| \leq \kappa I^2\) for all 
    \(\theta \in [0,2\pi]\),
    \(r \in \, ]\underline{r},\overline{r}[\),
    and \(I \in [-I_2,I_2]\).
    Taking this fact alongside the assumption that \(S(r_0) \geq \epsilon\),
    we find a lower bound \(r_{n+1} \geq r_n + I\epsilon - \kappa I^2\).
    Let \(0 < I^\star < \min\{I_2, \epsilon/\kappa\}\).
    Then for each \(I \in ]0,I^\star]\), the value 
    \(\gamma := I\epsilon - \kappa I^2\) is strictly positive, so
    \(P(r_0) \geq r_0 + \gamma\).
    Therefore, \(r_{n} \geq r_0 + n\gamma\), 
    and all solutions of \eqref{eqn:r-discrete} eventually flow past
    \(\overline{r}\).

    We conclude that, for \(I \in \, ]0, I^\star]\), the constrained dynamics gain
    energy on \(D\).
    Changing the sign of \(I\) changes the direction of the
    above inequalites, so that \(r_{n} \leq r_0 - n\gamma\). 
    In this case all orbits of \eqref{eqn:r-discrete} flow below
    \(\underline{r}\), i.e., the constrained dynamics lose energy on \(D\).
\end{proof}

\subsection{Energy Gain on \(\mathcal{O}_1\)}\label{sec:proof-o1}
To prove the first part of Theorem
\ref{thm:acrobot-energy-stabilization}, we use
\cite[Eqn. (9)]{dynamic_vhcs_stabilize_closed_orbits} to find 
pseudo-polar coordinates adapted to level sets of \(E\) on
\(\mathcal{O}_1 \backslash \{(0,0)\}\).
The resulting coordinate transformation is
\begin{align}
    \label{eqn:r-osc}
    r &= \arccos\left(\cos(q_u) - \frac{p_u^2}{30m^2gl^3}\right)
    , \\
    \label{eqn:theta-osc}
    \theta &=
    \left.\arctan_2\left(-\sign{p_u}\sqrt{1-\frac{q_u^2}{r^2}},\frac{q_u}{r}\right)
    \right|_{r = r(q_u,p_u)}
    ,
\end{align}
where \((r,\theta) \in \, ]0,\pi[ \times \Sone\).
One can verify that the inverse to this transformation is
\begin{align*}
    q_u &= rc_\theta
    , \\
    p_u &= -\sign{s_\theta}
    \sqrt{30m^2gl^3\cos(rc_\theta - c_r)}
    ,
\end{align*}
where for shorthand we write \(c_\theta := \cos(\theta)\), \(s_\theta :=
\sin(\theta)\), \(c_r := \cos(r)\), and \(s_r := \sin(r)\).
We now verify these pseudo-polar coordinates satisfy the requirements of 
Lemma \ref{lemma:energy-regulation-D}.
We omit the expressions for \(f_r(\theta,r,I)\) and \(f_\theta(\theta,r,I)\), as
they are too long for print.

It is straightforward to show that
\(f_r(\theta,r,0) \equiv 0\)
and
\begin{equation} \label{eqn:theta-dot-nom-osc}
    f_\theta(\theta,r,0) = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(r c_\theta) - c_r}
            {r^2 s_\theta^2}}
    .
\end{equation}
Note that ~\eqref{eqn:theta-dot-nom-osc} has removable singularities at
\(\theta \in \{0,\pi\}\).
Taking the limit as \(\theta\) approaches these points yields
\begin{equation}\label{eqn:theta-dot-limit}
    \lim\limits_{\theta \to 0}f_\theta(\theta,r,0)
    = \lim\limits_{\theta \to \pi} f_\theta(\theta,r,0)
    = \sqrt{\frac{6g s_r}{10lr}}
    , 
\end{equation}
which is smooth and well-defined for all
\(r \in \, ]0,\pi[\).
From \eqref{eqn:theta-dot-nom-osc}--\eqref{eqn:theta-dot-limit},
one can verify that \(f_\theta(\theta,r,0) > 0\) as required.
Defining \(g(\theta,r,I)\) as in Lemma \ref{lemma:energy-regulation-D}, one can
also verify that \((\partial g/\partial r)(\theta,r,0) \equiv 0\).

Symbolic computations reveal that
\(b(\theta,r_0) = K a(\theta,r_0)\),
where
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    a(\theta,r_0) &:= \frac{
        r_0 |s_\theta| \left(
        5 c_{r_0} \cos(r_0 c_\theta) - 8 \cos(r_0c_\theta)^2 + 3
    \right)
    }{
    s_{r_0}\sqrt{\cos(r_0c_\theta) - c_{r_0}}
    }
    .
\end{align*}
Notice that \(K\) is a positive constant which depends only on
\(m\), \(g\), \(l\), and \(\bar{q}_a\), while \(a(\theta,r_0)\) is adimensional
and \(2\pi\)-periodic in \(\theta\).
This means the integral of \(a(\theta,r_0)\) can be computed once for all
acrobots; 
hence, there exists \(\epsilon > 0\) such that 
\(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
if and only if there exists \(\delta > 0\) such that 
\(Q(r_0) := \int_0^{2\pi} a(\sigma,r_0) d\sigma \geq \delta\)
for every \(r_0 \in ]0,\pi[\).

Unfortunately, this latter statement is not true.
The plot of \(Q(r_0)\) for \(r_0 \in [10^{-10}, \pi - 10^{-3}]\) is shown in
Figure \ref{fig:acrobot-Q}.
We see that it is strictly positive and monotonically increasing, with an
asymptote at \(r_0 = \pi\), but \(Q(r_0) \to 0\) as \(r_0 \to 0\).

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{Qro.png}
    \caption{The plot of \(Q(r_0)\).}
    \label{fig:acrobot-Q}
\end{figure}

Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
appropriate subset of \(\mathcal{O}_1\).
To find this subset, first recall the discrete-time pseudo-radius dynamics
\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\).
The origin is now a repeller, which means there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium for \eqref{eqn:r-discrete} instead.

Let \(E_{\leq\mu}\) be the sublevel set of the energy \(E(\mu,0)\).
Since there exists \(\delta_\mu > 0\) such that \(Q(r_0) \geq \delta_\mu\) 
for \(r_0 \in [\mu,\pi[\), all conditions of Lemma
\ref{lemma:energy-regulation-D} are valid on the annulus
\(D_\mu = \mathcal{O}_1 \backslash E_{\leq\mu}\). 
Hence, there exists some \(I_1(\mu) > 0\) such that the acrobot gains/loses
energy everywhere on \(D_\mu\).
Taking \(I^\star = \min\{I_0,I_1(\mu)\}\), the acrobot will gain energy on
\(\mathcal{O}_1\) for \(I \in \, ]0,I^\star]\) and lose energy 
for \(I \in [-I^\star,0[\).

\subsection{All acrobots will rotate}

We have proven that any acrobot constrained by 
our VNHC \eqref{eqn:acrobot-constraint} will gain (or lose) energy on
\(\mathcal{O}_1\).
We must now prove that every acrobot will begin rotating, 
i.e., that when \(I \in \, ]0,I^\star]\) almost every orbit initialized in
\(\mathcal{O}_1\) will reach the line \(|q_u| = \pi\).

The definition of energy gain only guarantees that orbits initialized in
\(\mathcal{O}_1\) will escape compact subsets of \(\mathcal{O}_1\) in finite
time. 
For an orbit of the acrobot to reach the line \(|q_u| = \pi\), it must first
escape the closure \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) through
its boundary, the level set \(E_\pi\) with energy level \(E(\pi,0)\).

Taking the Jacobian of the constrained dynamics
~\eqref{eqn:acrobot-constrained-dynamics} at the upright equilibrium 
\((q_u,p_u) = (\pi,0)\) yields
\[
    J = \begin{bmatrix}
        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
        3mgl & mgl\bar{q}_aI
    \end{bmatrix}
    ,
\]
which has characteristic polynomial
\[
    \det\left(\lambda \Id{2} - J\right)
    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
    .
\]
This polynomial always has exactly one root with positive real part. 
The equilibrium \((\pi,0)\) is therefore unstable, so the stable manifold
\(\Pi^+\) of initial conditions converging to \((\pi,0)\) is one-dimensional,
and hence is of measure zero in \(\SxR\).
